---
title: "Global warming and wine"
---

```{r}
#| label: setup
#| message: false

library(dplyr)
library(ggplot2)
library(ggiraph)
library(plotly)
library(crosstalk)

source("R/wrangle_wine.R")
source("R/wrangle_co2.R")

wine <- get_wine_data()
co2_country <- get_co2_data()
co2_world <- get_co2_world()

min_vintage <- min(wine$vintage, na.rm = TRUE)
max_co2_year <- max(co2_country$year)
```

## Cumulative CO2 emissions and wine quality, side by side

The slider below spans `r min_vintage`\u2013`r max_co2_year`: from the oldest
wine vintage in the dataset through the most recent year available in the
CO2 data. **Note**: vintages before roughly 2000 are extremely sparse in the
wine dataset (often a single wine per year), so the wine-quality map will
look mostly empty for early years — this is a data coverage limitation, not
a real absence of production.

```{r}
#| label: combined-data
#| message: false

wine_country_year <- wine |>
  filter(!is.na(vintage), !is.na(country_iso3)) |>
  group_by(country_iso3, vintage) |>
  summarise(
    avg_points = mean(points, na.rm = TRUE),
    n_high_90 = sum(points >= 90, na.rm = TRUE),
    n_wines = n(),
    .groups = "drop"
  )

combined <- co2_country |>
  filter(year >= min_vintage, year <= max_co2_year) |>
  select(country, iso_code, year, cumulative_co2) |>
  left_join(
    wine_country_year,
    by = c("iso_code" = "country_iso3", "year" = "vintage")
  ) |>
  mutate(
    co2_hover = paste0("<b>", country, "</b><br>", year, "<br>Cumulative CO2: ", round(cumulative_co2, 0), " Mt"),
    wine_hover = ifelse(
      is.na(avg_points),
      paste0("<b>", country, "</b><br>", year, "<br>No wine reviews for this vintage"),
      paste0(
        "<b>", country, "</b><br>", year,
        "<br>Avg. score: ", round(avg_points, 1),
        "<br>Wines scoring \u226590: ", n_high_90,
        " / ", n_wines
      )
    )
  )

shared_combined <- SharedData$new(combined, key = ~iso_code, group = "co2_year")
```

::: {.panel-tabset}

### Cumulative CO2 emissions

```{r}
#| label: co2-map
#| message: false

plot_geo(shared_combined) |>
  add_trace(
    z = ~cumulative_co2, color = ~cumulative_co2, colors = "OrRd",
    locations = ~iso_code, text = ~co2_hover, hoverinfo = "text"
  ) |>
  colorbar(title = "Cumulative\nCO2 (Mt)") |>
  layout(geo = list(showframe = FALSE, showcoastlines = FALSE, projection = list(type = "natural earth")))
```

### Average wine score

```{r}
#| label: wine-map
#| message: false

plot_geo(shared_combined) |>
  add_trace(
    z = ~avg_points, color = ~avg_points, colors = "RdYlGn",
    locations = ~iso_code, text = ~wine_hover, hoverinfo = "text"
  ) |>
  colorbar(title = "Avg. score\n(pts)") |>
  layout(geo = list(showframe = FALSE, showcoastlines = FALSE, projection = list(type = "natural earth")))
```

:::

```{r}
#| label: year-slider

filter_slider("year", "Year", shared_combined, ~year, step = 1, width = "100%")
```

Use the tabs to switch between the two maps and the slider above to move
through years — both maps are linked to the same year selection, so you can
flip back and forth to compare a given year's cumulative emissions against
that vintage's wine scores. Bear in mind this is a purely descriptive,
country-level comparison: wine quality here reflects only the country's
average review score for wines reviewed by Wine Enthusiast in that vintage,
which is affected by which countries/producers happened to be reviewed, not
a controlled measure of climate impact on quality.

## Does wine-growing latitude shift as CO2 accumulates?

As a proxy for whether high-quality wine production is shifting toward
higher latitudes as the climate warms, we track the (unweighted) mean
absolute latitude of each vintage's highly-rated wines (points \u2265 90)
against global cumulative CO2 emissions for that year. We restrict to years
with at least 10 such wines, since earlier/sparser years would otherwise be
dominated by one or two data points.

```{r}
#| label: latitude-co2
#| message: false
#| fig-height: 5

wine_year_lat <- wine |>
  filter(!is.na(vintage), points >= 90, !is.na(wine_region_latitude)) |>
  group_by(vintage) |>
  summarise(mean_abs_latitude = mean(abs(wine_region_latitude)), n_wines = n(), .groups = "drop")

lat_co2 <- wine_year_lat |>
  inner_join(co2_world, by = c("vintage" = "year")) |>
  filter(n_wines >= 10)

spearman_lat <- cor(lat_co2$mean_abs_latitude, lat_co2$cumulative_co2, method = "spearman")

girafe(ggobj = ggplot(lat_co2, aes(cumulative_co2, mean_abs_latitude)) +
  geom_point_interactive(aes(tooltip = paste0(vintage, ": ", round(mean_abs_latitude, 1), "\u00b0, n=", n_wines), size = n_wines)) +
  geom_path(alpha = 0.3) +
  labs(
    x = "Global cumulative CO2 emissions (Mt)",
    y = "Mean |latitude| of \u226590-point wines",
    size = "N wines"
  ))
```

Spearman correlation: `r round(spearman_lat, 2)`.

Over the `r min(lat_co2$vintage)`\u2013`r max(lat_co2$vintage)` window covered
by high-scoring wines with known region latitude, the correlation between
global cumulative CO2 and mean production latitude is weak
(`r round(spearman_lat, 2)`). This is not strong evidence of a
climate-driven shift: the wine-region latitude here is a coarse,
country-level approximation (not per-vineyard), the time window is short
and dominated by 1990s–2010s data, and which countries/producers get
reviewed each year likely changes for reasons unrelated to climate (market
growth, reviewer assignment, publication coverage). Treat this as a
starting point for a much more careful analysis (e.g. region-level
coordinates, controlling for which countries are reviewed each year) rather
than a conclusive finding.
