---
title: "Global warming and wine"
---

```{r}
#| label: setup
#| message: false

library(dplyr)
library(tidyverse)
library(plotly)
library(crosstalk)
library(highcharter)
library(countrycode)
library(RColorBrewer)

source("R/wrangle_wine.R")
source("R/wrangle_co2.R")

wine <- get_wine_data()
co2_country <- get_co2_data()
co2_world <- get_co2_world()

min_vintage <- min(wine$vintage, na.rm = TRUE)
max_co2_year <- max(co2_country$year)
```

## Cumulative CO2 emissions and wine quality

The two maps below are stacked so both are always visible. The **cumulative
CO2** map is static — it shows each country's all-time cumulative emissions
up until 2021, since cumulative totals only grow monotonically and
aren't meaningfully "sliced" by a single vintage year. The slider underneath
filters only the **wine score** map, letting you step through vintages
1904-2021. **Note**: vintages before roughly 2000
are extremely sparse in the wine dataset (often a single wine per year), so
the wine-quality map will look mostly empty/uniform for early years — this is
a data coverage limitation, not a real absence of production.

```{r}
#| label: co2-map
#| message: false

co2_latest <- co2_country |>
  filter(year == max_co2_year) |>
  transmute(iso_code, country, value = cumulative_co2)

hcmap(
  "custom/world",
  data = co2_latest,
  joinBy = c("iso-a3", "iso_code"),
  value = "value",
  name = "Cumulative CO2 (Mt)",
  borderWidth = 0.1
) |>
  hc_colorAxis(minColor = "#fff7ec", maxColor = "#7f0000", type = "logarithmic") |>
  hc_title(text = paste0("Cumulative CO2 emissions by country (as of ", max_co2_year, ")")) |>
  hc_tooltip(pointFormat = "{point.name}: {point.value:,.0f} Mt") |>
  hc_mapNavigation(enabled = TRUE)
```

```{r}
#| label: wine-year-data
#| message: false

wine_years <- wine |>
  filter(
    !is.na(vintage), !is.na(country_iso3),
    vintage >= min_vintage, vintage <= max_co2_year
  ) |>
  group_by(country_iso3, vintage) |>
  summarise(
    value = round(mean(points, na.rm = TRUE), 1),
    n_high_90 = sum(points >= 90, na.rm = TRUE),
    n_wines = n(),
    .groups = "drop"
  ) |>
  transmute(
    iso_code = country_iso3, year = vintage, value, n_wines, n_high_90,
    key = paste(iso_code, year, sep = "_"),
    hover = paste0(
      iso_code, ", ", year, "<br>Avg. score: ", value,
      "<br>Wines scoring \u226590: ", n_high_90, " / ", n_wines
    )
  )

shared_wine <- SharedData$new(wine_years, key = ~key, group = "wine_year_filter")

# Every ISO3 country code, so the map can show the full world (countries
# without a wine review for the selected year(s) are drawn in flat grey
# rather than left blank).
world_countries <- countrycode::codelist |>
  filter(!is.na(iso3c)) |>
  distinct(iso3c) |>
  transmute(iso_code = iso3c)

# Fixed across the *entire* wine_years range (not the currently-filtered
# subset), so a given score always maps to the same color regardless of
# which year(s) the slider has selected.
wine_value_range <- range(wine_years$value, na.rm = TRUE)
```

```{r}
#| label: wine-map
#| message: false

# plotly has built-in crosstalk support, so the map filters directly from the
# slider below with no manual JS/aggregation needed. `value` is already a
# plain (unweighted) per-country/vintage average. The grey trace is drawn
# first (and isn't a crosstalk SharedData, so it always shows every
# country); the wine-score trace draws on top and only overrides countries
# that have data for the current filter.

ylgn_colors <- brewer.pal(9, "YlGn")  # -color light yellow -> dark green ramp
ylgn_colorscale <- Map(function(x, col) list(x, col),
                        seq(0, 1, length.out = length(ylgn_colors)),
                        ylgn_colors)


plot_geo() |>
  add_trace( # default view with all countries
    data = world_countries,
    locations = ~iso_code, locationmode = "ISO-3",
    z = 1, colorscale = list(list(0, "#e0e0e0"), list(1, "#e0e0e0")),
    showscale = FALSE, hoverinfo = "text",
    text = "No wine reviews for the selected year(s)"
  ) |>
  add_trace( # scored view
    data = shared_wine,
    locations = ~iso_code, locationmode = "ISO-3",
    z = ~value, zauto = FALSE, zmin = wine_value_range[1], 
                              zmax = wine_value_range[2],
    colorscale = ylgn_colorscale,
    text = ~hover, hoverinfo = "text"
  ) |>
  colorbar(title = "Avg. score\n(pts)") |>
  layout(
    title = "Average wine score by country",
    geo = list(showframe = FALSE, showcoastlines = FALSE, 
    projection = list(type = "natural earth"))
  )
```

```{r}
#| label: year-slider

filter_slider("year_filter", "Year", shared_wine, ~year, step = 1, width = "100%")
```

Drag the slider to move through vintages — only the wine-score map above
responds. If you select a range spanning multiple years, the map will show
individual country/vintage rows for each year in that range rather than a
blended value. Bear in mind this is a purely descriptive, country-level
comparison: wine quality here reflects only the country's average review
score for wines reviewed by Wine Enthusiast in that vintage, which is
affected by which countries/producers happened to be reviewed, not a
controlled measure of climate impact on quality.

## Does wine-growing latitude shift as CO2 accumulates?

As a proxy for whether high-quality wine production is shifting toward
higher latitudes as the climate warms, we track the (unweighted) mean
absolute latitude of each vintage's highly-rated wines (points \u2265 90)
against global cumulative CO2 emissions for that year. We restrict to years
with at least 10 such wines, since earlier/sparser years would otherwise be
dominated by one or two data points.

```{r}
#| label: latitude-co2
#| message: false
#| fig-height: 5

wine_year_lat <- wine |>
  filter(!is.na(vintage), points >= 90, !is.na(wine_region_latitude)) |>
  group_by(vintage) |>
  summarise(mean_abs_latitude = mean(abs(wine_region_latitude)), n_wines = n(), .groups = "drop")

lat_co2 <- wine_year_lat |>
  inner_join(co2_world, by = c("vintage" = "year")) |>
  filter(n_wines >= 10) |>
  arrange(vintage)

spearman_lat <- cor(lat_co2$mean_abs_latitude, lat_co2$cumulative_co2, method = "spearman")

plot_ly(lat_co2) |>
  add_lines(
    x = ~vintage, y = ~cumulative_co2, name = "Cumulative CO2 (Mt)",
    line = list(color = "firebrick"), yaxis = "y"
  ) |>
  add_markers(
    x = ~vintage, y = ~mean_abs_latitude, name = "Mean |latitude| of \u226590-pt wines",
    yaxis = "y2",
    marker = list(
      size = ~n_wines, sizeref = 2 * max(lat_co2$n_wines) / (40^2), sizemode = "area",
      color = "steelblue"
    ),
    text = ~paste0(vintage, ": ", round(mean_abs_latitude, 1), "\u00b0, n=", n_wines),
    hoverinfo = "text"
  ) |>
  layout(
    xaxis = list(title = "Vintage year"),
    yaxis = list(title = "Global cumulative CO2 (Mt)"),
    yaxis2 = list(title = "Mean |latitude| (\u00b0)", overlaying = "y", side = "right"),
    legend = list(orientation = "h", y = -0.2)
  )
```

Spearman correlation: `r round(spearman_lat, 2)`.

Over the `r min(lat_co2$vintage)`\u2013`r max(lat_co2$vintage)` window covered
by high-scoring wines with known region latitude, the correlation between
global cumulative CO2 and mean production latitude is weak
(`r round(spearman_lat, 2)`). This is not strong evidence of a
climate-driven shift: the wine-region latitude here is a coarse,
country-level approximation (not per-vineyard), the time window is short
and dominated by 1990s–2010s data, and which countries/producers get
reviewed each year likely changes for reasons unrelated to climate (market
growth, reviewer assignment, publication coverage). Treat this as a
starting point for a much more careful analysis (e.g. region-level
coordinates, controlling for which countries are reviewed each year) rather
than a conclusive finding.
