---
title: "Wine quality in the world"
---

```{r}
#| label: setup
#| message: false

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggiraph)
library(tidytext)
library(highcharter)
library(magrittr)
library(readr)
library(glmnet)
library(gt)
```

This page explores the `winemag` wine-tasting reviews: how price varies by
province, whether points predict price, whether reviewers specialize, and
what language distinguishes highly- vs. poorly-rated wines.

## Which province has the highest average price?

```{r}
#| label: fig-price
#| fig-height: 5

wine <- read_csv("winemag-data-130k-v2.csv")
colnames(wine)
data(GNI2014, package = "treemap") # This data is needed to get the iso3  from the country name

prov_price <- wine |>
  filter(!is.na(province), !is.na(price)) |>
  group_by(country) |>
  summarise(avg_price = mean(price), n_wines = n(), .groups = "drop") |>
  left_join(GNI2014 |> dplyr::select(iso3, country) |> distinct(), by = "country")
```

```{r}
#| label: fig-province-price
#| fig-height: 5
hcmap(map = "custom/world-robinson-lowres", 
  download_map_data = getOption("highcharter.download_map_data"), 
  data = prov_price,
  value = "avg_price", 
  joinBy = c("iso-a3", "iso3")) |> 
    hc_colorAxis(stops = color_stops(colors = viridisLite::inferno(10, begin = 0.1)))
```
@fig-province-price shows the average price by country.

```{r}
#| label: fig-nwines-price
#| fig-height: 5
hcmap(map = "custom/world-robinson-lowres", 
  download_map_data = getOption("highcharter.download_map_data"), 
  data = prov_price,
  value = "n_wines", 
  joinBy = c("iso-a3", "iso3")) |> 
    hc_colorAxis(stops = color_stops(colors = viridisLite::inferno(10, begin = 0.1)))
```
@fig-nwines-price shows the number of wines by country.

```{r}
#| label: fig-province-price-density
#| fig-height: 5

prov_price %$% 
  hchart(avg_price, color = "#17b8b6", name = "average price") |> 
  hc_title(text = "Distribution of average price by country")
```
@fig-province-price-density shows the distribution of average price by country.

```{r}
#| label: fig-province-nwines-density
#| fig-height: 5

prov_price %$% 
  hchart(n_wines, color = "#17b8b6", name = "number of wines") |> 
  hc_title(text = "Distribution of the number of wines by country")
```
@fig-province-nwines-density shows the distribution of the number of wines by country.


## Do points predict price?

Price is strongly right-skewed (a few wines cost hundreds of dollars while
most are under $50), so a plain Pearson correlation on raw price is
misleading. We compare **Spearman's rank correlation** (robust to skew) with
Pearson's correlation computed on **log-price**.

```{r}
#| label: points-price-corr

wp <- wine |> filter(!is.na(price), !is.na(points))
spearman_r <- cor(wp$points, wp$price, method = "spearman")

# With a single predictor, lasso shrinkage has no variables to select
# between, so glmnet() refuses to fit; a simple linear model on log-price
# gives the equivalent Pearson correlation on that scale.
pearson_raw_r <- cor(wp$points, wp$price)
```
- Spearman rank correlation is: `r round(spearman_r, 2)`


```{r}
log_price_model <- lm(log(price) ~ points, data = wp)
pearson_r_logprice <- cor(wp$points, log(wp$price))
summary(log_price_model) 
```
- Pearson correlation on log-price is: `r round(pearson_r_logprice, 2)`
- The log-price model explains `r round(summary(log_price_model)$r.squared, 2)` of the variation in log-price.


```{r}
#| label: fig-points-vs-price
#| fig-height: 5
#| message: false
#| cache: true

hchart(wp, "scatter", hcaes(x = points, y = log2(price+1)))
```
@fig-points-vs-price shows the relationship between points and log-price.


