---
title: "Wine quality in the world"
---

```{r}
#| label: setup
#| message: false

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggiraph)
library(tidytext)
library(highcharter)
library(magrittr)
```

This page explores the `winemag` wine-tasting reviews: how price varies by
province, whether points predict price, whether reviewers specialize, and
what language distinguishes highly- vs. poorly-rated wines.

## Which province has the highest average price?

```{r}
#| label: province-price
#| fig-height: 5

wine <- read_csv("winemag-data-130k-v2.csv")
colnames(wine)
data(GNI2014, package = "treemap") # This data is needed to get the iso3  from the country name

prov_price <- wine |>
  filter(!is.na(province), !is.na(price)) |>
  group_by(country) |>
  summarise(avg_price = mean(price), n_wines = n(), .groups = "drop") |>
  left_join(GNI2014 |> dplyr::select(iso3, country) |> distinct(), by = "country")


hcmap(map = "custom/world-robinson-lowres", 
  download_map_data = getOption("highcharter.download_map_data"), 
  data = prov_price,
  value = "avg_price", 
  joinBy = c("iso-a3", "iso3")) |> 
    hc_colorAxis(stops = color_stops(colors = viridisLite::inferno(10, begin = 0.1)))


prov_price %$% plot(density(avg_price))
prov_price %$% 
  hchart(avg_price, color = "#17b8b6", name = "average price") |> 
  hc_title(text = "Distribution of average price by country")

```

Unfiltered, the top provinces by average price are dominated by
provinces with only a handful of wines (e.g. `r top_all$province[1]` has
just `r top_all$n_wines[1]` review(s)) — a small-sample artifact rather than
a meaningful signal.

```{r}
#| label: province-price-filtered
#| fig-height: 5

girafe(ggobj = plot_province_price(top_10plus, "Provinces with \u226510 wines"))
```

Restricting to provinces with at least 10 reviews, `r top_10plus$province[1]`
has the highest average price at $`r round(top_10plus$avg_price[1], 0)`,
a more credible ranking than the unfiltered version.

## Do points predict price?

Price is strongly right-skewed (a few wines cost hundreds of dollars while
most are under $50), so a plain Pearson correlation on raw price is
misleading. We compare **Spearman's rank correlation** (robust to skew) with
Pearson's correlation computed on **log-price**.

```{r}
#| label: points-price-corr

wp <- wine |> filter(!is.na(price), !is.na(points))

spearman_r <- cor(wp$points, wp$price, method = "spearman")
pearson_log_r <- cor(wp$points, log(wp$price), method = "pearson")
pearson_raw_r <- cor(wp$points, wp$price, method = "pearson")
```

- Pearson correlation on raw price: `r round(pearson_raw_r, 2)`
- Pearson correlation on log(price): `r round(pearson_log_r, 2)`
- Spearman rank correlation: `r round(spearman_r, 2)`

```{r}
#| label: points-price-plot
#| fig-height: 5
#| message: false

set.seed(4831)
wp_sample <- wp |> slice_sample(n = 6000)

girafe(ggobj = ggplot(wp_sample, aes(points, price)) +
  geom_jitter_interactive(
    aes(tooltip = paste0(title, "\n$", price, ", ", points, " pts")),
    alpha = 0.3, width = 0.3, height = 0
  ) +
  scale_y_log10() +
  labs(x = "Points", y = "Price (USD, log scale)"))
```

The raw-price Pearson correlation (`r round(pearson_raw_r, 2)`) understates
the relationship because of price's skew; both the log-price Pearson
correlation and the Spearman correlation agree on a **moderate positive**
association (~0.6) — higher-scoring wines tend to cost more, but points
explain only part of the variation in price (plenty of highly-rated wines
remain affordable, and some expensive wines score modestly).

## Reviewer-level trends: do reviewers specialize?

For each reviewer we compute a concentration index (Herfindahl–Hirschman
Index, HHI) over the provinces and varieties they've reviewed. HHI ranges
from close to 0 (reviews spread evenly across many categories) to 1 (all
reviews concentrated in a single category).

```{r}
#| label: reviewer-specialization
#| fig-height: 6

specialization <- function(df, group_col) {
  df |>
    filter(!is.na(taster_name), !is.na({{ group_col }})) |>
    count(taster_name, {{ group_col }}) |>
    group_by(taster_name) |>
    mutate(total = sum(n), share = n / total) |>
    summarise(
      n_reviews = sum(n),
      n_categories = n(),
      hhi = sum(share^2),
      .groups = "drop"
    )
}

spec_province <- specialization(wine, province) |> mutate(dimension = "Province")
spec_variety <- specialization(wine, variety) |> mutate(dimension = "Variety")

spec_combined <- bind_rows(spec_province, spec_variety) |>
  filter(n_reviews >= 50)

girafe(ggobj = ggplot(spec_combined, aes(hhi, reorder(taster_name, hhi), fill = dimension)) +
  geom_col_interactive(
    aes(tooltip = paste0(taster_name, " (", dimension, "): HHI=", round(hhi, 2), ", ", n_categories, " categories, n=", n_reviews)),
    position = "dodge"
  ) +
  labs(x = "Concentration (HHI)", y = NULL, fill = NULL))
```

Concentration varies widely across reviewers. Some (e.g. reviewers with very
few total reviews or working for a single publication's regional beat) are
highly concentrated in one or two provinces or varieties (HHI near 1),
while others such as Roger Voss or Michael Schachner cover dozens of
provinces and varieties with low concentration — consistent with broad,
generalist coverage rather than specialization. Note that very low review
counts inflate apparent specialization, so this chart already filters to
reviewers with at least 50 reviews.

## What language distinguishes positive and negative reviews?

We define **positive** reviews as points \u2265 90 and **negative** reviews as
points \u2264 85 (roughly the top and bottom quartiles). Rather than comparing
raw term frequency — which mostly surfaces generic wine vocabulary common to
both groups ("fruit", "palate", "finish") — we compute a **log-odds ratio**
of each word's frequency in positive vs. negative reviews, keeping only
words that appear at least 200 times across both groups combined.

```{r}
#| label: text-mining
#| fig-height: 6
#| message: false

data("stop_words")

wine_words <- wine |>
  filter(review_sentiment %in% c("positive", "negative")) |>
  select(review_sentiment, description) |>
  unnest_tokens(word, description) |>
  anti_join(stop_words, by = "word") |>
  filter(!word %in% c("wine", "wines"), !str_detect(word, "^\\d+$"))

word_freq <- wine_words |>
  count(review_sentiment, word) |>
  pivot_wider(names_from = review_sentiment, values_from = n, values_fill = 0) |>
  filter(positive + negative >= 200) |>
  mutate(
    prop_pos = (positive + 1) / (sum(positive) + 1),
    prop_neg = (negative + 1) / (sum(negative) + 1),
    log_ratio = log2(prop_pos / prop_neg)
  )

top_terms <- bind_rows(
  word_freq |> slice_max(log_ratio, n = 15) |> mutate(sentiment = "More common in positive reviews"),
  word_freq |> slice_min(log_ratio, n = 15) |> mutate(sentiment = "More common in negative reviews")
)

ggplot(top_terms, aes(log_ratio, reorder(word, log_ratio))) +
  geom_col() +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Log2 odds ratio (positive vs. negative)", y = NULL)
```

Positive reviews lean on terms like "gorgeous", "stunning", and "wonderful",
along with structural/aging descriptors ("grained" tannins, notions of a
"decade" of aging potential). Negative reviews cluster around flaws and
plainness: "watery", "bland", "generic", "dull", "thin", "flat" — vocabulary
describing a lack of concentration or character rather than an outright
technical fault.
