---
title: "Wine quality in the world"
---

```{r}
#| label: setup
#| message: false

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggiraph)
library(tidytext)
library(highcharter)
library(magrittr)
library(readr)
library(glmnet)
library(gt)
```


## Wine price by country
```{r}
#| label: fig-price
#| fig-height: 5

wine <- read_csv("winemag-data-130k-v2.csv")
colnames(wine)
data(GNI2014, package = "treemap") # This data is needed to get the iso3  from the country name

prov_price <- wine |>
  filter(!is.na(province), !is.na(price)) |>
  mutate(country = ifelse(country=="US", "United States", country)) |>
  mutate(country = ifelse(country=="Macedonia", "Macedonia, FYR", country)) |>
  mutate(country = ifelse(country=="Slovakia", "Slovak Republic", country)) |>
  mutate(country = ifelse(country=="England", "United Kingdom", country)) |>
  group_by(country) |>
  summarise(avg_price = mean(price), n_wines = n(), .groups = "drop") |>
  left_join(GNI2014 |> dplyr::select(iso3, country) |> distinct(), by = "country")
```

### Top 5 countries by average wine price
```{r}
#| column: margin

prov_price |> 
  arrange(desc(avg_price)) |> head(5) |>
  hchart(
    type = "column",
    hcaes(x = country, y = avg_price),
    options3d = list(enabled = TRUE, beta = 15, alpha = 15))
```

## Number of wines by country
```{r}
#| label: fig-province-price
#| fig-height: 5
hcmap(map = "custom/world-robinson-lowres", 
  download_map_data = getOption("highcharter.download_map_data"), 
  data = prov_price,
  value = "avg_price", 
  joinBy = c("iso-a3", "iso3")) |> 
    hc_colorAxis(stops = color_stops(colors = viridisLite::inferno(10, begin = 0.1)))
```
@fig-province-price shows the average price by country.

### Top 5 countries by number of wines
```{r}
#| column: margin

prov_price |> 
  arrange(desc(n_wines)) |> head(5) |>
  hchart(
    type = "column",
    hcaes(x = country, y = n_wines),
    options3d = list(enabled = TRUE, beta = 15, alpha = 15))
```

## Distribution of wine prices by country
```{r}
#| label: fig-nwines-price
#| fig-height: 5
hcmap(map = "custom/world-robinson-lowres", 
  download_map_data = getOption("highcharter.download_map_data"), 
  data = prov_price,
  value = "n_wines", 
  joinBy = c("iso-a3", "iso3")) |> 
    hc_colorAxis(stops = color_stops(colors = viridisLite::inferno(10, begin = 0.1)))
```
@fig-nwines-price shows the number of wines by country.


## Distribution of average price and number of wines by country
```{r}
#| label: fig-province-price-density
#| fig-height: 5

prov_price %$% 
  hchart(avg_price, color = "#17b8b6", name = "average price") |> 
  hc_title(text = "Distribution of average price by country")
```
@fig-province-price-density shows the distribution of average price by country.

### Distribution of number of wines by country
```{r}
#| label: fig-province-nwines-density
#| fig-height: 5

prov_price %$% 
  hchart(n_wines, color = "#17b8b6", name = "number of wines") |> 
  hc_title(text = "Distribution of the number of wines by country")
```
@fig-province-nwines-density shows the distribution of the number of wines by country.


## Do points predict price?

### Correlation between points and price
```{r}
#| label: points-price-corr

wp <- wine |> filter(!is.na(price), !is.na(points))
spearman_r <- cor(wp$points, wp$price, method = "spearman")

# With a single predictor, lasso shrinkage has no variables to select
# between, so glmnet() refuses to fit; a simple linear model on log-price
# gives the equivalent Pearson correlation on that scale.
pearson_raw_r <- cor(wp$points, wp$price)
```
- Spearman rank correlation is: `r round(spearman_r, 2)`

### linear model if points predict log-price
```{r}
log_price_model <- lm(log2(price+1) ~ points, data = wp)
pearson_r_logprice <- cor(wp$points, log2(wp$price+1))
summary(log_price_model) 
```
- Pearson correlation on log-price is: `r round(pearson_r_logprice, 2)`
- The log-price model explains `r round(summary(log_price_model)$r.squared, 2)` of the variation in log-price.

## Relationship between points and log-price
```{r}
#| label: fig-points-vs-price
#| fig-height: 5
#| message: false
#| cache: true

hchart(wp, "scatter", hcaes(x = points, y = log2(price+1)))
```
@fig-points-vs-price shows the relationship between points and log-price.

