---
title: "Wine quality in the world"
---

```{r}
#| label: setup
#| message: false

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggiraph)
library(tidytext)
library(highcharter)
library(magrittr)
library(readr)
```

This page explores the `winemag` wine-tasting reviews: how price varies by
province, whether points predict price, whether reviewers specialize, and
what language distinguishes highly- vs. poorly-rated wines.

## Which province has the highest average price?

```{r}
#| label: fig-price
#| fig-height: 5

wine <- read_csv("winemag-data-130k-v2.csv")
colnames(wine)
data(GNI2014, package = "treemap") # This data is needed to get the iso3  from the country name

prov_price <- wine |>
  filter(!is.na(province), !is.na(price)) |>
  group_by(country) |>
  summarise(avg_price = mean(price), n_wines = n(), .groups = "drop") |>
  left_join(GNI2014 |> dplyr::select(iso3, country) |> distinct(), by = "country")
```

```{r}
#| label: fig-province-price
#| fig-height: 5
hcmap(map = "custom/world-robinson-lowres", 
  download_map_data = getOption("highcharter.download_map_data"), 
  data = prov_price,
  value = "avg_price", 
  joinBy = c("iso-a3", "iso3")) |> 
    hc_colorAxis(stops = color_stops(colors = viridisLite::inferno(10, begin = 0.1)))
```
@fig-province-price shows the average price by country.

```{r}
#| label: fig-nwines-price
#| fig-height: 5
hcmap(map = "custom/world-robinson-lowres", 
  download_map_data = getOption("highcharter.download_map_data"), 
  data = prov_price,
  value = "n_wines", 
  joinBy = c("iso-a3", "iso3")) |> 
    hc_colorAxis(stops = color_stops(colors = viridisLite::inferno(10, begin = 0.1)))
```
@fig-nwines-price shows the number of wines by country.

```{r}
#| label: fig-province-price-density
#| fig-height: 5

prov_price %$% 
  hchart(avg_price, color = "#17b8b6", name = "average price") |> 
  hc_title(text = "Distribution of average price by country")
```
@fig-province-price-density shows the distribution of average price by country.

```{r}
#| label: fig-province-nwines-density
#| fig-height: 5

prov_price %$% 
  hchart(n_wines, color = "#17b8b6", name = "number of wines") |> 
  hc_title(text = "Distribution of the number of wines by country")
```
@fig-province-nwines-density shows the distribution of the number of wines by country.
